Fast and scalable Gaussian process modeling of stellar variability \vspace{-3\baselineskip}

### 2016-12-18 paper/ms.tex
Matched notation

@@

The general ACF is given by
\begin{equation}
K(\tau) = a_r e^{-b_r \vert\tau\vert} \cos{(b_r \vert\tau\vert)} + a_i e^{-b_r \vert\tau\vert} \sin{(b_r \vert\tau\vert)}.
\end{equation}
which has a power spectrum given by
\begin{equation}
P(\omega) = P_0 \frac{(a_ib_i+a_rb_r)(b_i^2+b_r^2)+(a_rb_r-a_ib_i)\omega^2}{(\omega^2-(b_i^2+b_r^2))^2 + 4b_r^2\omega^2}.
\end{equation}
The power spectrum should be non-negative since it is the absolute square of the Fourier transform of the noise.
Note that the denominator is always non-negative, so the power spectrum is non-negative if $a_ib_i+a_rb_r \ge 0$ and
$a_rb_r-a_ib_i \ge 0$.

@@

This power spectrum matches that of our basis kernel if $\omega_0^2 = b_r^2+b_i^2$,
$a_rb_r-a_ib_i = 0$, $a_ib_i+a_rb_r=\omega_0^2$, and $4b_r^2 = \omega_0^2/Q^2$.
The power spectrum can be normalized by $P_0$ in amplitude and $\omega_0$ in
frequency to yield a spectral shape that only depends upon $Q$.  This is plotted
in Figure \ref{sho_psd} for various values of $Q$.  This power spectrum has
a peak frequency given by $\omega_{max} = \omega_0\sqrt{1-1/(2Q^2)}$, which is
greater than zero for $Q > 1/\sqrt{2}$.

@@

This power spectrum has several limits of interest:
\begin{itemize}
\item $Q=1/2$:  This results in a power spectrum $\propto 1/(x^2+1)^2$, where $x=\omega/\omega_0$.
This matches the power spectrum of the Mat\'ern 3/2 kernel, and can be represented by the following
autocovariance function:
\begin{equation}
K(\tau) = a_r e^{-b_r \vert\tau\vert} \left(\cos{(b_i \vert\tau\vert)} + \frac{b_r}{b_i}\sin{(b_i \vert\tau\vert)}\right),
\end{equation}
in the limit that $b_i \rightarrow \infty$, for which the second term has a coefficients that diverges,
but with large values of $b_i$, this may approximate the Mat\'ern kernel well.  Note that this
power spectrum peaks at $\omega=0$.
\item $Q=1/\sqrt{2}$:  This results in a power spectrum $\propto 1/(x^4+1)$, which is a functional
form used to fit the Solar granulation power spectrum by \cite{2009A&A...495..979M}.  This kernel has
dependence
\begin{equation}
K(\tau) = a_r e^{-b_r\vert\tau\vert} \left(\cos{(b_i\vert\tau\vert)}+\sin{(b_i\vert\tau\vert)}\right).
\end{equation}
Despite the sinusoidal dependence of the kernel, the strong damping causes the power spectrum to
peak at zero frequency; such a low quality oscillator doesn't display even quasi-periodic behavior.
\item $Q \gg 1$: This represents a high-quality oscillator which approaches a narrow peak.  The
power spectrum approaches
\begin{equation}
P(\omega) \approx P_0 \frac{Q^2}{1-1/(4Q^2)+4x_{max}^2\left(x-x_{max}\right)^2Q^2},
\end{equation}
where $x_{max}=\sqrt{1-1/(2Q^2)}$.  This shape approaches a Lorentzian for $x \approx x_{max}$ with
a width of $1/(2Qx_{max})$ and a peak height $\propto Q^2$.
\end{itemize}

>>>>

Parameters:
\begin{eqnarray}\eqlabel{kernel}
k(\tau) &=& \sum_{j=1}^p \left[
    a_j\,\exp\left(-c_j\,\tau\right)\,\cos\left(-d_j\,\tau\right) +
    b_j\,\exp\left(-c_j\,\tau\right)\,\sin\left(-d_j\,\tau\right)
\right]
\end{eqnarray}
where $\tau = |\Delta t|$.
Equivalently, this can be written as
\begin{eqnarray}
k(\tau) &=& \sum_{j=1}^p \left[
    \frac{1}{2}(a_j + i\,b_j)\,\exp\left(-(c_j+i\,d_j)\,\tau\right) +
    \frac{1}{2}(a_j - i\,b_j)\,\exp\left(-(c_j-i\,d_j)\,\tau\right)
\right]\quad.
\end{eqnarray}

The power spectral density is computed by taking the Fourier transform of
\eq{kernel} to find
\begin{eqnarray}
S(\omega) = \sum_{j=1}^p \sqrt{\frac{2}{\pi}}
\frac{(a_j\,c_j+b_j\,d_j)\,({c_j}^2+{d_j}^2)+(a_j\,c_j-b_j\,d_j)\,\omega^2}
{\omega^4+2\,({c_j}^2-{d_j}^2)\,\omega^2+({c_j}^2+{d_j}^2)^2}\quad.
\end{eqnarray}

For some $b_j=0$, the PSD for the $j$-th component becomes
\begin{eqnarray}
S_j(\omega) = \frac{1}{\sqrt{2\,\pi}}\,\frac{a_j}{c_j}\,\left[
    \frac{1}{1+\left(\frac{\omega-d_j}{c_j}\right)^2} +
    \frac{1}{1+\left(\frac{\omega+d_j}{c_j}\right)^2}
\right]
\end{eqnarray}
a pair of Lorentzian distributions with width $c_j$ centered on
$\omega = \pm d_j$.

@@

%The general ACF is given by
%\begin{equation}
%K(\tau) = a_r e^{-b_r \vert\tau\vert} \cos{(b_r \vert\tau\vert)} + a_i e^{-b_r \vert\tau\vert} \sin{(b_r \vert\tau\vert)}.
%\end{equation}
%which has a power spectrum given by
%\begin{equation}
%P(\omega) = P_0 \frac{(a_ib_i+a_rb_r)(b_i^2+b_r^2)+(a_rb_r-a_ib_i)\omega^2}{(\omega^2-(b_i^2+b_r^2))^2 + 4b_r^2\omega^2}.
%\end{equation}
%The power spectrum should be non-negative since it is the absolute square of the Fourier transform of the noise.
%Note that the denominator is always non-negative, so the power spectrum is non-negative if $a_ib_i+a_rb_r \ge 0$ and
%$a_rb_r-a_ib_i \ge 0$.

@@

% Check to see that signs match
This power spectrum matches that of our basis kernel if $\omega_0^2 = c_j^2+d_j^2$,
$a_jc_j-b_jd_j = 0$, $b_jd_j+a_jc_j=\omega_0^2$, and $4c_j^2 = \omega_0^2/Q^2$.
The power spectrum can be normalized by $P_0$ in amplitude and $\omega_0$ in
frequency to yield a spectral shape that only depends upon $Q$.  This is plotted
in Figure \ref{sho_psd} for various values of $Q$.  This power spectrum has
a peak frequency given by $\omega_{max} = \omega_0\sqrt{1-1/(2Q^2)}$, which is
greater than zero for $Q > 1/\sqrt{2}$.

@@

This power spectrum has several limits of interest:
\begin{itemize}
\item $Q=1/2$:  This results in a power spectrum $\propto 1/(x^2+1)^2$, where $x=\omega/\omega_0$.
This matches the power spectrum of the Mat\'ern 3/2 kernel, and can be represented by the following
autocovariance function:
\begin{equation}
K(\tau) = a_j e^{-c_j \vert\tau\vert} \left(\cos{(d_j \vert\tau\vert)} + \frac{c_j}{d_j}\sin{(d_j \vert\tau\vert)}\right),
% Check to see that signs match in this equation.
\end{equation}
in the limit that $b_i \rightarrow \infty$, for which the second term has a coefficients that diverges,
but with large values of $b_i$, this may approximate the Mat\'ern kernel well.  Note that this
power spectrum peaks at $\omega=0$.
\item $Q=1/\sqrt{2}$:  This results in a power spectrum $\propto 1/(x^4+1)$, which is a functional
form used to fit the Solar granulation power spectrum by \cite{2009A&A...495..979M}.  This kernel has
dependence
\begin{equation}
K(\tau) = a_r e^{-b_r\vert\tau\vert} \left(\cos{(b_i\vert\tau\vert)}+\sin{(b_i\vert\tau\vert)}\right).
\end{equation}
Despite the sinusoidal dependence of the kernel, the strong damping causes the power spectrum to
peak at zero frequency; such a low quality oscillator doesn't display even quasi-periodic behavior.
\item $Q \gg 1$: This represents a high-quality oscillator which approaches a narrow peak.  The
power spectrum approaches
\begin{equation}
P(\omega) \approx P_0 \frac{Q^2}{1-1/(4Q^2)+4x_{max}^2\left(x-x_{max}\right)^2Q^2},
\end{equation}
where $x_{max}=\sqrt{1-1/(2Q^2)}$.  This shape approaches a Lorentzian for $x \approx x_{max}$ with
a width of $1/(2Qx_{max})$ and a peak height $\propto Q^2$.
\end{itemize}
